---
title: "hw1"
author: "Yifeng Huang"
format: typst
editor: visual
---

This is the link to the repo: https://github.com/YifengHuang-Bob/Homework1


# Problem 1 - Exploratory Analysis of Soybean Cultivars

## a.

```{r}
soybean <- read.csv("data/data.csv")

names(soybean) <- c(
  "Season", "Cultivar", "Repetition", "PH", "IFP", "NLP",
  "NGP", "NGL", "NS", "MHG", "GY"
)

class(soybean)
names(soybean)
```
\

## b.

```{r}
nrow(soybean)

length(unique(soybean$Cultivar))

length(unique(soybean$Season))

table(soybean$Season)

```
The dataset contains 320 observations, 40 distinct cultivars, and observations from 2 seasons. There are 160 observations from Season 1 and 160 observations from Season 2.

\

## c.

### 1.

```{r}
cor(soybean$PH, soybean$GY)
```
The correlation between plant height (PH) and grain yield (GY) is approximately 0.123.


### 2.

```{r}
cor(
  soybean$PH[soybean$Season == 1],
  soybean$GY[soybean$Season == 1]
)

cor(
  soybean$PH[soybean$Season == 2],
  soybean$GY[soybean$Season == 2]
)
```
The correlation between PH and GY is approximately 0.231 for Season 1 and -0.023 for Season 2. Therefore, Season 1 has the higher correlation.


### 3.

```{r}
soybean$GY[which.max(soybean$PH)]
```
The grain yield of the soybean observation with the greatest plant height is 3304.678.


### 4.

```{r}
mean_mhg <- mean(soybean$MHG)

mean_mhg

mean(soybean$MHG > mean_mhg) * 100
```
47.1875% of the observations have an MHG greater than the overall average MHG.


### 5.

```{r}
mean_gy_by_cultivar <- tapply(
  soybean$GY,
  soybean$Cultivar,
  mean
)

mean_gy_by_cultivar[which.max(mean_gy_by_cultivar)]
```
The cultivar with the highest average grain yield is 98R30 CE, with an average grain yield of 3964.968.


\


## d.

```{r}
quantitative_vars <- c(
  "PH", "IFP", "NLP", "NGP", "NGL", "NS", "MHG", "GY"
)

average_table <- rbind(
  Overall = colMeans(soybean[quantitative_vars]),
  `Season 1` = colMeans(
    soybean[soybean$Season == 1, quantitative_vars]
  ),
  `Season 2` = colMeans(
    soybean[soybean$Season == 2, quantitative_vars]
  )
)

round(average_table, 2)
```

\

## e.

### 1.

Let $\mu_1$ be the mean grain yield (GY) for Season 1 and $\mu_2$ be the mean grain yield for Season 2.

$H_0: \mu_1 = \mu_2$

$H_A: \mu_1 \neq \mu_2$


### 2.

```{r}
t.test(
  soybean$GY[soybean$Season == 1],
  soybean$GY[soybean$Season == 2]
)
```

#### Manual calculation

```{r}
gy1 <- soybean$GY[soybean$Season == 1]
gy2 <- soybean$GY[soybean$Season == 2]

n1 <- length(gy1)
n2 <- length(gy2)

mean1 <- mean(gy1)
mean2 <- mean(gy2)

var1 <- var(gy1)
var2 <- var(gy2)

t_stat <- (mean1 - mean2) /
  sqrt(var1 / n1 + var2 / n2)

df <- (var1 / n1 + var2 / n2)^2 /
  ((var1 / n1)^2 / (n1 - 1) +
     (var2 / n2)^2 / (n2 - 1))

p_value <- 2 * pt(-abs(t_stat), df = df)

t_stat
df
p_value

```


The mean grain yield is 3428.452 for Season 1 and 3408.655 for Season 2. The Welch two-sample t-test gives a p-value of 0.7255. Since the p-value is greater than 0.05, we fail to reject the null hypothesis. There is insufficient evidence to conclude that the mean grain yield differs between the two seasons.

\

# Problem 2 - Adult Income Dataset

## a.

```{r}
adult <- read.csv(
  "data/adult.data",
  header = FALSE,
  strip.white = TRUE,
  na.strings = "?"
)

class(adult)
dim(adult)
```

\

## b.

```{r}
names(adult) <- c(
  "age", "workclass", "fnlwgt", "education", "education_num",
  "marital_status", "occupation", "relationship", "race", "sex",
  "capital_gain", "capital_loss", "hours_per_week",
  "native_country", "income"
)

names(adult)
```

\

## c.

```{r}
colSums(is.na(adult))
```
Missing values occur only in `workclass`, `occupation`, and `native_country`. There are 1836 missing observations for `workclass`, 1843 for `occupation`, and 583 for `native_country`. All other variables have no missing observations.

\

## d.

```{r}
nrow(adult)

adult_filtered <- adult[adult$hours_per_week >= 10, ]

nrow(adult_filtered)
```
The sample size is 32,561 before filtering and 32,103 after restricting the sample to individuals working at least 10 hours per week.

\

## e.

```{r}
nrow(adult_filtered)

adult_complete <- adult_filtered[
  !is.na(adult_filtered$occupation) &
    !is.na(adult_filtered$workclass) &
    !is.na(adult_filtered$native_country),
]

nrow(adult_complete)

colSums(
  is.na(
    adult_complete[c("occupation", "workclass", "native_country")]
  )
)
```
Observations with missing values in occupation, workclass, or native_country were removed because these variables are needed for the analysis, and observations with missing values in these fields would have incomplete information.

\

## f.

```{r}
adult_clean <- adult_complete[
  adult_complete$age >= 18 &
    adult_complete$age <= 90 &
    adult_complete$education != "Preschool",
]

range(adult_clean$age)
sum(adult_clean$education == "Preschool")
```

\

## g.

```{r}
nrow(adult_clean)
```
The final sample size is 29,523 observations.















